---
title: "Compiling annual fishing fleet effort data"
format:
    html:
        fig-width: 10
        fig-height: 5
        code-fold: true
        dpi: 400
        toc: true
        toc-depth: 2
        embed-resources: true
---

```{r setup and load data, include=FALSE}
packages <- c("ggplot2", "arrow", "tidyverse", "glue", "tabulapdf", "gganimate", "randomcoloR")
sapply(packages, library, character.only = TRUE)
options(dplyr.summarise.inform = FALSE)

habitats <- readRDS("./Objects/Habitats.rds")
```

# Midwater Trawl
Effort data for the Midwater Trawl fishing fleet comes from Reed et al. (2007). This fleet mainly consists of one large trawler vessel performing the majority of midwater trawls in South Africa.
The midwater trawl fishery extends east from 22 degrees East. This means that most of the fleet's effort likely doesn't occur within the Southern Benguela model habitat domain. To get a more accurate estimate of effort we need to scale the total effort estimate by the proportion of effort occurring within the domain area.
```{r}
mw_intensity <- "../../Spatial Data/fishing_effort_data/Midwater_Trawl_Intensity/Midwater_Trawl_Intensity/Midwater_Trawl_Intensity.tif"
mw_intensity <- project(rast(mw_intensity), "epsg:4326", method = "near")
mw_intensity <- as.numeric(mw_intensity)

total_spatial_mw_intensity <- global(mw_intensity, fun = "sum", na.rm = TRUE)
prop_mw_in_domain <- exact_extract(mw_intensity, habitats, fun = "sum")
prop_mw_in_domain <- sum(prop_mw_in_domain) / total_spatial_mw_intensity

annual_mw_effort <- 5092.3 * prop_mw_in_domain[1,1] # Number of trawls estimated from Reed et al. 2007. scaled by the proportion of effort in habitat domain.
```

# Nets including small scale

# Linefishery

# Squid jig

# Longlines

# Purse Seine

# Demersal Trawl

# West Coast Rock Lobster traps
