---
title: "Choosing StrathE2E domain extent - South Africa"
author: "Jack H Laverick, University of Strathclyde"
date: "`r format(Sys.time(), '%A %d %b')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = "center")

Tidy_packages <- c("tidyverse", "data.table", "pbapply", "furrr", "viridis", "patchwork") # List handy data packages
Geo_packages <- c("sf", "rnaturalearth",  "raster", "stars")                # List GIS packages
lapply(c(Tidy_packages, Geo_packages), library, character.only = TRUE)      # Load packages
source("./R scripts/@_Region file.R")                                       # Define project region 

plan(multisession)                                                          # Choose the method to parallelise by

domain <- readRDS("./Objects/Domains.rds")

world <- ne_countries(scale = "medium", returnclass = "sf") %>%             # Get a world map
  st_transform(crs = crs)                                                   # Assign polar projection


```

StrathE2E contains 3 ocean volumes. There's an inshore-offshore divide, and a shallow-deep divide. Only the offshore volume is split into shallow and deep. Choosing a shallow-deep boundary depth has knock-on consequences for the inshore-offshore divide, as the inshore zone will be defined as the sea surface to this depth. StrathE2E assumes that the volumes are well mixed, and any exchange between the deep and shallow compartments are controlled by vertical diffusivity at that depth. The boundary depth therefore needs to be deep enough to avoid excessive mixing between the volumes, while also considering how large the inshore zone will end up being. The spatial arrangement of StratheE2E is shown below.

![](../Shared data/spatial schematic.svg)

For the Benguela current we have inspected the distribution of vertical diffusivity values in the region from NEMO-MEDUSA model output. 50 m looks like a sensible depth to separate the shallow and deep layers. **And aligns with other models in the region**.

As well as the vertical boundary depth, We need to choose the horizontal extent for the inshore and offshore zones of the model domain. When doing this we need to consider the following: 

* The offshore zone should extend to the continental slope. 
* Fishing activity is often high along the shelf edge, so this activity should be within in the model domain. 
* The depth limits for the inshore zone will be 0 and the minimum depth for the offshore zone.
* The inshore zone must be an area large enough to sample NEMO-MEDUSA model outputs.

In StrathE2E terrestrial inputs can only reach the offshore zone by transiting through the inshore zone. In areas with very steep bathymetry it is possible the offshore zone polygons created above could contact land. We therefore also need to set a minimum distance from shore for the offshore zone. This ensures the inshore zone exists all along the coast. In practice this distance needs to allow a pixel of NEMO-MEDUSA model output to sit between the coast and offshore zone polygon. We chose a minimum distance from shore for the offshore zone of 20 km, which typically captures the data we need.

We have suggested a 500 m maximum depth for the offshore zone as a start point, to track the contour of the continental shelf.
**Updated to 800 m to capture Hake trawls.**

Finally we need to "cut" the polygons defined by depths and distances from shore so that they don't extend around the entire landmass. As a first suggestion we have limited the model domain to the North where the continental shelf constricts after the orange river cell (near the Namibian border). To the South we limit to where the Agulhas bank constricts. If there are reasons to choose different breaks, please let us know.

This results in the following:

```{r domain, eval = T}

colours <- c(Inshore = "yellow", Offshore = "yellow3")

 GEBCO <- read_stars("../Shared data/GEBCO_2020.nc", proxy = TRUE)
 st_crs(GEBCO) <- 4326

 crop <- domain %>% 
   st_buffer(40000) %>% 
   st_transform(4326) %>% 
   st_bbox()
 
 GEBCO <- GEBCO[crop] %>%
   st_transform(crs) %>%
   st_as_stars(downsample = 3)

land <- units::set_units(0, "m")

GEBCO[[1]][GEBCO[[1]] >= land] <- NA

ggplot() +
  geom_stars(data = GEBCO) +
  viridis::scale_fill_viridis(name = "Elevation", option = "turbo") +
  ggnewscale::new_scale_fill() +
  geom_sf(data = domain, aes(fill = Shore), colour = "black", size = 0.1) +
  #geom_sf(data = Region_mask, colour = "white", fill = NA) +
  geom_sf(data = world, size = 0.1, fill = "black") +
  scale_fill_manual(values = colours, name = "Zone") +
  zoom +
  theme_minimal() +
  labs(caption = "Suggested model area") +
  NULL
```

The polygons for the inshore and offshore zones are shown in yellows. The bathymetry is plotted underneath in a rainbow palette, so you can see how the shelf extends beyond where we cut the domain. 

To check how the proposed model domain relates to fishing activity, we plotted the cumulative fishing effort according to global fishing watch from 2017 - 2020:

``` {r fishing, eval = TRUE}

GFW <- readRDS("./Notes/Cache/domain_GFW.rds") %>% 
  st_transform(crs = crs)

ggplot() +
  geom_sf(data = world, size = 0.1, fill = "grey30", colour = "white", size = 0.1) +
  geom_sf(data = domain, fill = "grey", colour = "black", size = 0.1) +
  geom_stars(data = GFW, colour = NA) +
  geom_sf(data = domain, fill = NA, colour = "white", size = 0.2) +
  viridis::scale_fill_viridis(na.value = NA, trans = "log10", option = "inferno") +
  zoom +
  theme_minimal() +
  labs(caption = "Total fishing effort") +
  NULL

```

The model domain is shown in grey with white borders, while the fishing effort is shown as a heatmap on a log scale. They grey area appears to capture the majority of the fishing effort along the continental shelf. There is some fishing effort just outside the continental shelf which is not captured by our model domain. Any ideas on what this is likely to be? Pelagic or Demersal? And do you want the model domain adjusted to capture it?


``` {r Deep convection cached, eval = F}

b <- readRDS("./Notes/Cache/domain_V_b.rds")

b
```

``` {r Vertical diffusivity cached, eval = F}

c <- readRDS("./Notes/Cache/domain_V_c.rds")

c
```
